---
4title: DDS Final
author: "Keith Sikes"
date: "2022-08-03"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r}
install.packages("caret")
install.packages("rlang")
library(tidyverse)
library(rlang)
library(caret)
library(klaR)
```




```{r}
setwd("~/Personal/Rcoding/Final Project")

data_i <- read_csv("CaseStudy2-data.csv")
```

For this data set we wanted to pull out everyone who was 55 or over to show if they were quitting, retiring, or staying with the company.

```{r}

d_retired = data_i %>% 
  mutate(Retired = ifelse(Attrition == "Yes" & Age >=55, "Yes", "No") ) %>% 
  mutate(Attrition_Detail = ifelse(Retired == "Yes", "Retired", ifelse(Attrition == "Yes", "Quit", "Stayed") ) ) %>% 
  group_by(Attrition_Detail) %>% summarise(Count = n()) %>% 
  mutate(Pct = round(Count / sum(Count),2))

```

In these charts we pulled out several of the key factors that related to attrition. We seperated data out of each column, created a mean of attrition, and showed an average for how many stayed vs the amount left.The last two charts show the ages that people left and the results from type of attrition.

```{r}

to_convert = c("Education","EnvironmentSatisfaction","JobInvolvement","JobLevel","JobSatisfaction","NumCompaniesWorked","PerformanceRating","RelationshipSatisfaction","StockOptionLevel","TrainingTimesLastYear","WorkLifeBalance","BusinessTravel")
data = data_i %>% 
  mutate(across(all_of(to_convert), as.character, .names = "{col}_CAT" )) %>%  
  filter(! (Attrition == "Yes" & Age >=55) )  


for(col in colnames(data)){
  plot_list = list()
  if(typeof(data[[col]]) == "character" & col != "Attrition" ){
    d_group = data %>% group_by(!!sym(col) ) %>% 
      mutate(Attrition_TF = Attrition == "Yes") %>% 
      summarise(Mean = mean(Attrition_TF), Count = n() )
    plot = ggplot(d_group, aes(!!sym(col), Mean)) + geom_col() + ggtitle(col) +
      geom_text(aes(label = Count), nudge_y = .01) +
  theme(plot.title = element_text(hjust = 0.5))
    print(plot)
  }
}

ggplot(data = data_i, aes(x = Age)) + 
  geom_histogram(color="snow4", fill="#8B8989") +
  ggtitle("Age of Attrition") +
  xlab("Age") +
  ylab("Quantity") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75)) +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data = d_retired, aes(x = Count, y = Attrition_Detail )) + 
  geom_col(color="snow4", fill="#8B8989") + coord_flip() +
  ggtitle("Types of Attrition") +
  xlab("Attrition") +
  ylab("Count") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75)) +
  theme(plot.title = element_text(hjust = 0.5))

```

In this Naive Bayes confustion matrix model we show the amount of people to expect to leave vs those that will stay within an 78% accuracy.

```{r}

#use numcompanieswork, jobrole, overtime
#make the data for the ML model
vars_for_model = c("OverTime","IsSatisfied","IsSalesRep","IsInvolved","IsHigherJobLevel","IsStockOption_0_3","WorkLifeBalance")
d_ml = data %>% 
  mutate(IsSatisfied = EnvironmentSatisfaction > 1) %>% 
  mutate(IsSalesRep = (JobRole == "Sales Representative") ) %>% 
  mutate(IsInvolved = JobInvolvement > 1) %>% 
  mutate(IsHigherJobLevel = JobLevel > 1) %>% 
  mutate(IsStockOption_0_3 = StockOptionLevel %in% c(0,3)) %>% 
  mutate(WorkLifeBalance = WorkLifeBalance > 1) %>% 
  mutate(across(.cols = everything(), as.factor)) %>% 
  mutate(Attrition = factor(Attrition, levels = c("Yes","No"))) %>% 
  dplyr::select(Attrition, all_of(vars_for_model))
rownames(d_ml) <- NULL

library(caTools)
library(e1071)

#break into training and test sets(without caret)
set.seed(101) 
trainIndex = sample.split(d_ml$Attrition, SplitRatio = .7)
d_train <- d_ml[ trainIndex,]
d_test  <- d_ml[-trainIndex,]

#calculate ratio between yes's and no's
num_yes = d_train %>% filter(Attrition=="Yes") %>% nrow()
num_no = nrow(d_train) - num_yes
ratio = round(num_no / num_yes)

#generate the needed # of samples to add yes's
to_add = tibble()
for(i in 1:(ratio-0) ){
  to_add = bind_rows(to_add, d_train %>% filter(Attrition=="Yes") ) 
}

#add rows to d_ml
d_train = d_train %>% bind_rows(to_add)

#Make a dataframe from X and Y
X_train = d_train %>% dplyr::select(-Attrition)
y_train = d_train %>% dplyr::select(Attrition)

#
X_test = d_test %>% dplyr::select(-Attrition)
y_test = d_test %>% dplyr::select(Attrition)

#
model = train(X_train,y_train$Attrition,'nb') #,trControl=trainControl(method='cv',number=10)

#confusion matrix
conf_mat = table(predict(model$finalModel,X_test)$class,y_test$Attrition)
conf_mat_pct = prop.table(conf_mat) %>% round(2)
conf_mat
conf_mat_pct

# calculate sensitity
pred = predict(model$finalModel,X_test)$class
truth = y_test$Attrition
sensitivity(pred, truth)
specificity(pred, truth)
```


```{r}
#FOR TRAINING SET

#confusion matrix
conf_mat = table(predict(model$finalModel,X_train)$class,y_train$Attrition)
conf_mat_pct = prop.table(conf_mat) %>% round(2)
conf_mat
conf_mat_pct

# calculate sensitity
pred = predict(model$finalModel,X_train)$class
truth = y_train$Attrition
sensitivity(pred, truth)
specificity(pred, truth)

```
Now we have to import the data without attrition and predict if they will leave or stay
```{r}
noattdata <- read_csv("https://raw.githubusercontent.com/sikesmaurice/sikesmaurice.github.io/main/CaseStudy2CompSet%2BNo%2BAttrition.csv")

#vars_for_model = c("OverTime","IsSatisfied","IsSalesRep","IsInvolved","IsHigherJobLevel","IsStockOption_0_3","WorkLifeBalance")
noattdatamutate = noattdata %>% 
  mutate(IsSatisfied = EnvironmentSatisfaction > 1) %>% 
  mutate(IsSalesRep = (JobRole == "Sales Representative") ) %>% 
  mutate(IsInvolved = JobInvolvement > 1) %>% 
  mutate(IsHigherJobLevel = JobLevel > 1) %>% 
  mutate(IsStockOption_0_3 = StockOptionLevel %in% c(0,3)) %>% 
  mutate(WorkLifeBalance = WorkLifeBalance > 1) %>% 
  mutate(across(.cols = everything(), as.factor))
rownames(noattdatamutate) <- NULL

noattdatamutate = noattdatamutate[ , c("OverTime","IsSatisfied","IsSalesRep","IsInvolved","IsHigherJobLevel","IsStockOption_0_3","WorkLifeBalance")]

noattdatamutate$Attrition <- predict(model$finalModel, noattdatamutate)$class

NoAtExport <- data.frame(ID = noattdata$ID, Attrition = noattdatamutate$Attrition)

write.csv(NoAtExport, file = file.choose(new = T), row.names = FALSE)

```

# Salary Prediction

What factors in the data have an effect on salary

```{r}

data <- read_csv("https://raw.githubusercontent.com/sikesmaurice/sikesmaurice.github.io/main/CaseStudy2-data.csv")

# Too Many Unique Variables - Business Travel, Education Field, Department, Job Role, 

# IncomeNumberData <- select(data, -c("BusinessTravel", "EducationField", "Department", "JobRole", "DistanceFromHome", "ID", "Over18", "StandardHours", "MaritalStatus", "EmployeeNumber", "EmployeeCount"))

DirtyIncomeData <- which(colnames(data) %in% c("BusinessTravel", "EducationField", "Department", "JobRole", "DistanceFromHome", "ID", "Over18", "StandardHours", "MaritalStatus", "EmployeeNumber", "EmployeeCount"))

IncomeNumberData <- data[,-DirtyIncomeData]

IncomeNumberData$Gender <- as.factor(IncomeNumberData$Gender)
IncomeNumberData$Attrition <- as.factor(IncomeNumberData$Attrition)
IncomeNumberData$OverTime <- as.factor(IncomeNumberData$OverTime)

IncomeNumberData$Gender <- as.numeric(IncomeNumberData$Gender)
IncomeNumberData$Attrition <- as.numeric(IncomeNumberData$Attrition)
IncomeNumberData$OverTime <- as.numeric(IncomeNumberData$OverTime)

CorMat <- cor(IncomeNumberData)
MonInCor <- CorMat[,"MonthlyIncome"] # Shows relationship to Monthly Income Value (The closer to zero, the less correlation).
MonInCor[MonInCor>0.3]
# MonInCor[MonInCor<-0.3] Nothing was less than -0.3

# Variables to include: "Age", "JobLevel", "MonthlyIncome", "TotalWorkingYears", "YearsAtCompany", "YearsInCurrentRole", "YearsSinceLastPromotion", "YearsWithCurrManager"

```
# Making the Linear Regression Model to Predict Salary

In this Linear Regression Model we are able to show that the RMSE is $1273.571
```{r}

library(Metrics)

#Splitting the Data Set into Training and Test

set.seed(900)
Salary_Train <- createDataPartition(IncomeNumberData$MonthlyIncome, p = .8, list = FALSE, times = 1)
S_train <- IncomeNumberData[ Salary_Train,]
S_test  <- IncomeNumberData[-Salary_Train,]
S_cl <- S_test$MonthlyIncome

Sx_Test <- which(colnames(S_test) %in% c("MonthlyIncome"))
Sy_test1 <- S_test[,-Sx_Test]

# SalaryModel1 <- lm(MonthlyIncome ~ Age + JobLevel + TotalWorkingYears + YearsAtCompany + YearsInCurrentRole + YearsWithCurrManager, data = IncomeNumberData)

# Remove factors without stars by pvalues

SalaryModel <- lm(MonthlyIncome ~ JobLevel + TotalWorkingYears + YearsWithCurrManager, data = S_train)

SalaryPred <- predict(SalaryModel,Sy_test1)
rmse(SalaryPred, S_test$MonthlyIncome)

summary(SalaryModel)

```
```{R}

NoSalary <- read_csv("https://raw.githubusercontent.com/sikesmaurice/sikesmaurice.github.io/main/CaseStudy2CompSet%2BNo%2BSalary.csv")

NoSalaryFiltered <- NoSalary[,c("JobLevel", "TotalWorkingYears", "YearsWithCurrManager")]

NoSalaryPred <- predict(SalaryModel, NoSalaryFiltered)

NoSalaryExport <- data.frame(ID = NoSalary$ID, MonthlySalary = NoSalaryPred)

write.csv(NoSalaryExport, file = file.choose(new = T), row.names = FALSE)

```
# Visualizing the Data

Since we now know 

```{R}

NoSalary <- read_csv("https://raw.githubusercontent.com/BStrzalkowski/Data-Science-Project-2-/main/CaseStudy2CompSetNoSalary.csv")

NoSalaryFiltered <- NoSalary[,c("JobLevel", "TotalWorkingYears", "YearsWithCurrManager")]

NoSalaryHat <- predict(SalaryModel, NoSalaryFiltered)

NoSalaryExport <- data.frame(ID = NoSalary$ID, MonthlySalary = NoSalaryHat)

write.csv(NoSalaryExport, file = file.choose(new = T), row.names = FALSE)

```

